Evolution of solar wind discontinuities in the inner heliosphere

### 2025-06-12 notebooks/000_data.qmd
update

Interactively exploring the quantities of solar wind.

@@

@@

## Setup

```{julia}
using Dates
using DrWatson
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using Discontinuity
using SPEDAS
```

## Overview

```{julia}
using Speasy
using SPEDAS: ulabel, smooth, tnorm_combine, tplot
using DimensionalData
using CairoMakie
# using GLMakie
using Statistics
using PartialFunctions
using Unitful
using Beforerr: _theme_legend, easy_save
using SPEDAS.SpaceDataModel: set, extend, @set
include("../src/psp_temp.jl")
include("../src/calc.jl")
include("../src/meta.jl")

update_theme!(_theme_legend(; padding=(0, 0, 0, 0)))
```

```{julia}
# along time get the temperature from anisotropic temperature
function tTemp(x; dims=Ti)
    return mean.(eachslice(x; dims))
end

mva_transform((V, B), tmin, tmax; tb_min=tmin, tb_max=tmax) = mva(V(tmin, tmax), B(tb_min, tb_max))
make_mva_transform(tb_min, tb_max) = mva_transform$(; tb_min, tb_max)

j_transform((V, B), tmin, tmax) = current_density(B(tmin, tmax), V(tmin, tmax))
```

```{julia}
tmin_psp = DateTime("2021-01-14")
tmax_psp = DateTime("2021-01-20")
tmin_wind = DateTime("2021-01-15")
tmax_wind = DateTime("2021-01-23")
_tmax_wind = DateTime("2021-01-15T02")

tmin_psp = DateTime("2021-01-16")
tmax_psp = DateTime("2021-01-19")
# tmin_wind = DateTime("2021-01-17")
# tmax_wind = DateTime("2021-01-18")
```

```{julia}
tvars_psp = [
    set(tnorm_combine ∘ PSP.B_1MIN; ylabel=B_ylabel, labels=B_RTN_labels),
    set(PSP.n; ylabel=n_ylabel),
    set(PSP.V; ylabel=V_ylabel, labels=V_RTN_labels),
    set(PSP.T; ylabel=T_ylabel)
]
```

```{julia}
tvars_wind = [
    set(tnorm_combine ∘ Wind.B_GSE; ylabel=B_ylabel, labels=B_XYZ_labels),
    set(Wind.n; ylabel=n_ylabel),
    # set(Wind.V_GSE_p; ylabel=V_ylabel),
    replace_outliers! ∘ set(Wind.V_GSE_K0; ylabel=V_ylabel, labels=V_XYZ_labels),
    set(Wind.T; ylabel=T_ylabel)
]
```

```{julia}
using CairoMakie
CairoMakie.activate!()
```

```{julia}
using Intervals: Interval

function get_themis_tmask(t0, t1) 
    thx_v_gse = THEMIS.V_GSE(t0, t1)
    _times = @views SPEDAS.times(thx_v_gse)[(thx_v_gse[:, 1].>-200u"km/s").|(thx_v_gse[:, 2].<-50u"km/s")]
    dt = Minute(10)
    its = union([Interval(t - dt, t + dt) for t in _times])
    return tmask!(its)
end
```

```{julia}
tvars_themis = [
    set(tnorm_combine ∘ THEMIS.B_GSE; ylabel=B_ylabel, labels=B_XYZ_labels),
    set(THEMIS.n; ylabel=n_ylabel),
    replace_outliers! ∘ set(THEMIS.V_GSE; ylabel=V_ylabel, labels=V_XYZ_labels),
    set(THEMIS.eTemp; ylabel=T_ylabel)
]

```

```{julia}
for enc in (7, 8, 9)
    (tmin_psp, tmax_psp), (tmin_wind, tmax_wind) = get_timerange(enc)

    _tmask! = get_themis_tmask(tmin_wind, tmax_wind)

    legend = (; position=Top())
    labels = '1':'9'

    f = Figure(; size=(1200, 600))
    plots = tplot(f[1, 1], tvars_psp[1:3], tmin_psp, tmax_psp; legend)
    add_labels!(f[1, 1]; labels, open="(a.")
    tplot(f[1, 2], tvars_wind[1:3], tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 2]; labels, open="(b.")
    tplot(f[1, 3], _tmask! .∘ tvars_themis[1:3], tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 3]; labels, open="(c.")
    easy_save("overview-$enc")

    f = Figure(; size=(1200, 600))
    plots = tplot(f[1, 1], tvars_psp, tmin_psp, tmax_psp; legend)
    add_labels!(f[1, 1]; labels, open="(a.")
    tplot(f[1, 2], tvars_wind, tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 2]; labels, open="(b.")
    tplot(f[1, 3], _tmask! .∘ tvars_themis, tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 3]; labels, open="(c.")
    easy_save("overview-$enc-all")
end
```

## Event example (with pressure anisotropy)

```{julia}
using Discontinuity: ids_finder
tau = Second(30)
psp_events = ids_finder(PSP.B, tmin_psp, tmax_psp, tau)
wind_events = ids_finder(Wind.B_GSE, tmin_wind, tmax_wind, tau, Wind.V_GSE, Wind.n_p)
thm_events = ids_finder(THEMIS.B_FGL_GSE, tmin_wind, tmax_wind, tau)
```

```{julia}
psp_B_mva = set(mva ∘ PSP.B; name="mva_B", labels=B_mva_labels)
psp_V_mva = Product((PSP.V, PSP.B), mva_transform; ylabel=V_ylabel, labels=V_mva_labels)
psp_eTemp_ani = Product(read_electron_temperature(), tview; labels=["Electron (parallel)", "Electron (perpendicular)"])
psp_pTemp_ani = Product(read_proton_temperature(), tview; labels=["Proton (parallel)", "Proton (perpendicular)"])
psp_Temp_ani = [psp_eTemp_ani, psp_pTemp_ani]
psp_mva_all = [psp_B_mva, psp_V_mva, psp_Temp_ani]

thm_B_mva = set(mva ∘ THEMIS.B_FGL_GSE; name="mva_B", labels=B_mva_labels)
thm_V_mva = Product((THEMIS.V_GSE, THEMIS.B_FGL_GSE), mva_transform; ylabel=V_ylabel, labels=V_mva_labels)
thm_mva_all = [thm_B_mva, thm_V_mva, THEMIS.T_ani]
```

```{julia}
function make_mva_product(event, V, B=V; kwargs...)
    mva_transform = make_mva_transform(t_us_ds(event)...)
    return Product((V, B), mva_transform; kwargs...)
end

function make_mva_products(event, B, V, n=nothing, T3=nothing)
    B_mva = make_mva_product(event, B, B; labels=B_mva_labels, ylabel=B_ylabel)
    V_mva = make_mva_product(event, V, B; labels=V_mva_labels, ylabel=V_ylabel)
    J = Product((V_mva, B_mva), j_transform)
    result = [tnorm_combine ∘ B, V, tnorm_combine ∘ B_mva, set(tsubtract ∘ V_mva; ylabel=ΔV_ylabel), J]
    isnothing(n) || push!(result, set(
        tsubtract ∘ Product((B_mva, n), alfven_velocity_ts);
        ylabel=ΔV_ylabel, labels=["Va_L", "Va_M", "Va_N"]
    ))
    isnothing(T3) || begin
        Λ = Product((B_mva, n, T3), anisotropy_ts; ylabel="Λ")
        Va_Λ = set(
            tsubtract ∘ Product((B_mva, n, T3), alfven_velocity_ani_ts);
            ylabel=ΔV_ylabel, labels=["Va*_L", "Va*_M", "Va*_N"]
        )
        push!(result, Λ, Va_Λ)
    end
    return result
end
```

```{julia}
tvars2_wind = [
    tnorm_combine ∘ Wind.B,
    Wind.n,
    Wind.V_GSE,
    Wind.Temp,
    Wind.Temp_ani
]

tvars_themis2 = [[THEMIS.B_FGL_GSE, THEMIS.n, THEMIS.V_GSE, THEMIS.T]; thm_mva_all]
```

```{julia}
function plot_candidate(f, event, ts, toffset=Second(0); kwargs...)
    tmin, tmax = t_us_ds(event)
    f = tplot(f, ts, tmin - toffset, tmax + toffset; kwargs...)
    tlines!(f, [tmin, tmax])
    f
end

f = let
    f = Figure(; size=(1200, 800))
    # plot_candidate(f[1, 1], Event(psp_events, 2), [tvars_psp; psp_mva_all]; link_xaxes)

    wind_event = AniEvent(wind_events, 1) # 300
    wind_T3 = extend(Wind.pTemp_ani, Second(20))
    wind_tvars = make_mva_products(wind_event, Wind.B_GSE, Wind.V_GSE, Wind.n_p, wind_T3)
    wind_tvars = [wind_tvars[3:end]; Wind.Temp_ani; Wind.n; Wind.Temp]
    for p in Wind.Temp_ani
        @info "$p" data = p(wind_event)
    end
    wind_plots = plot_candidate(f, wind_event, wind_tvars, tau)
    # themis_plots = plot_candidate(f[1, 3], Event(thm_events, 3), tvars_themis2, Minute(1); link_xaxes)
    # hideydecorations!(wind_plots)
    # hideydecorations!(themis_plots)
end
```

## Batch processing

Wind Speed, Time to Earth
300 km/s, ~5.8 days
400 km/s, ~3.9 days
500 km/s, ~3.5 days
700 km/s, ~2.5 days

>>>>

@@

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SPEDAS: tplot
using GLMakie
```

### THEMIS

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        THEMIS.B_GSE,
        THEMIS.B_FGL_GSE,
        THEMIS.n,
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

### Wind

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        [Wind.V_GSE_3DP, Wind.V_GSE_K0],
        [Wind.n_p_K0, Wind.n_p_3DP, Wind.n_p_nonlin],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

`Wind.V_GSE_3DP` ("WI_PM_3DP/P_VELS") has higer time resolution than `Wind.V_GSE_K0` ("WI_K0_SWE/V_GSE").

`Wind.n_p_nonlin` and `Wind.n_p_K0` seem to be the same.

![](../figures/others/wind.png)

@@

### 2025-06-12 notebooks/00_index.qmd
update

Interactively exploring the quantities of solar wind.

## Data

- [PSP/FIELDS Level 2 and higher data](https://fields.ssl.berkeley.edu/data/), [SPDF - Coordinated Data Analysis Web (CDAWeb)](https://cdaweb.gsfc.nasa.gov/cgi-bin/eval2.cgi?dataset=PSP_FLD_L2_MAG_RTN&index=sp_phys)

WI_H1_SWE	Proton_W_moment	        Proton thermal speed W (km/s) from isotropic moment analysis
WI_H1_SWE	Proton_Wperp_moment 	Proton thermal speed Wperpendicular (km/s) from bimax moment analysis
WI_H1_SWE	Proton_Wpar_moment	    Proton thermal speed Wparallel (km/s) from bimax moment analysis

WI_H1_SWE	Proton_Wperp_nonlin	    Proton thermal speed perpendicular to the magnetic field direction [km/s].
WI_H1_SWE	Proton_Wpar_nonlin	    Proton thermal speed parallel to the magnetic field direction [km/s].
WI_PLSP_3DP	Ion omnidirectional fluxes 0.4-3 keV and moments, often at ~24 second resolution, PESA Low, Wind 3DP
WI_PM_3DP	Ion moments (computed on-board) @ 3 second (spin) resolution, PESA LOW, Wind 3DP

Notes: `thb_peim_t3_magQ` is not accurate in solar wind mode.

@@

```{julia}
using Dates
using DrWatson
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using Discontinuity
using SPEDAS
```

@@

```{julia}
using Speasy
using SPEDAS: ulabel, smooth, tnorm_combine, tplot
using DimensionalData
using CairoMakie
# using GLMakie
using Statistics
using PartialFunctions
using Unitful
using Beforerr: _theme_legend, easy_save
using SPEDAS.SpaceDataModel: set, extend, @set
include("../src/psp_temp.jl")
include("../src/calc.jl")
include("../src/meta.jl")

update_theme!(_theme_legend(; padding=(0, 0, 0, 0)))
```

@@

```{julia}
tvars_psp = [
    set(tnorm_combine ∘ PSP.B_1MIN; ylabel=B_ylabel, labels=B_RTN_labels),
    set(PSP.n; ylabel=n_ylabel),
    set(PSP.V; ylabel=V_ylabel, labels=V_RTN_labels),
    set(PSP.T; ylabel=T_ylabel)
]
```

```{julia}
tvars_wind = [
    set(tnorm_combine ∘ Wind.B_GSE; ylabel=B_ylabel, labels=B_XYZ_labels),
    set(Wind.n; ylabel=n_ylabel),
    # set(Wind.V_GSE_p; ylabel=V_ylabel),
    replace_outliers! ∘ set(Wind.V_GSE_K0; ylabel=V_ylabel, labels=V_XYZ_labels),
    set(Wind.T; ylabel=T_ylabel)
]
```

@@

```{julia}
tvars_themis = [
    set(tnorm_combine ∘ THEMIS.B_GSE; ylabel=B_ylabel, labels=B_XYZ_labels),
    set(THEMIS.n; ylabel=n_ylabel),
    replace_outliers! ∘ set(THEMIS.V_GSE; ylabel=V_ylabel, labels=V_XYZ_labels),
    set(THEMIS.eTemp; ylabel=T_ylabel)
]

```{julia}
for enc in (7, 8, 9)
    (tmin_psp, tmax_psp), (tmin_wind, tmax_wind) = get_timerange(enc)

@@

    legend = (; position=Top())
    labels = '1':'9'

    f = Figure(; size=(1200, 600))
    plots = tplot(f[1, 1], tvars_psp[1:3], tmin_psp, tmax_psp; legend)
    add_labels!(f[1, 1]; labels, open="(a.")
    tplot(f[1, 2], tvars_wind[1:3], tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 2]; labels, open="(b.")
    tplot(f[1, 3], _tmask! .∘ tvars_themis[1:3], tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 3]; labels, open="(c.")
    easy_save("overview-$enc")

    f = Figure(; size=(1200, 600))
    plots = tplot(f[1, 1], tvars_psp, tmin_psp, tmax_psp; legend)
    add_labels!(f[1, 1]; labels, open="(a.")
    tplot(f[1, 2], tvars_wind, tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 2]; labels, open="(b.")
    tplot(f[1, 3], _tmask! .∘ tvars_themis, tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 3]; labels, open="(c.")
    easy_save("overview-$enc-all")
end
```

@@

```{julia}
psp_B_mva = set(mva ∘ PSP.B; name="mva_B", labels=B_mva_labels)
psp_V_mva = Product((PSP.V, PSP.B), mva_transform; ylabel=V_ylabel, labels=V_mva_labels)
psp_eTemp_ani = Product(read_electron_temperature(), tview; labels=["Electron (parallel)", "Electron (perpendicular)"])
psp_pTemp_ani = Product(read_proton_temperature(), tview; labels=["Proton (parallel)", "Proton (perpendicular)"])
psp_Temp_ani = [psp_eTemp_ani, psp_pTemp_ani]
psp_mva_all = [psp_B_mva, psp_V_mva, psp_Temp_ani]

thm_B_mva = set(mva ∘ THEMIS.B_FGL_GSE; name="mva_B", labels=B_mva_labels)
thm_V_mva = Product((THEMIS.V_GSE, THEMIS.B_FGL_GSE), mva_transform; ylabel=V_ylabel, labels=V_mva_labels)
thm_mva_all = [thm_B_mva, thm_V_mva, THEMIS.T_ani]
```

```{julia}
function make_mva_product(event, V, B=V; kwargs...)
    mva_transform = make_mva_transform(t_us_ds(event)...)
    return Product((V, B), mva_transform; kwargs...)
end

function make_mva_products(event, B, V, n=nothing, T3=nothing)
    B_mva = make_mva_product(event, B, B; labels=B_mva_labels, ylabel=B_ylabel)
    V_mva = make_mva_product(event, V, B; labels=V_mva_labels, ylabel=V_ylabel)
    J = Product((V_mva, B_mva), j_transform)
    result = [tnorm_combine ∘ B, V, tnorm_combine ∘ B_mva, set(tsubtract ∘ V_mva; ylabel=ΔV_ylabel), J]
    isnothing(n) || push!(result, set(
        tsubtract ∘ Product((B_mva, n), alfven_velocity_ts);
        ylabel=ΔV_ylabel, labels=["Va_L", "Va_M", "Va_N"]
    ))
    isnothing(T3) || begin
        Λ = Product((B_mva, n, T3), anisotropy_ts; ylabel="Λ")
        Va_Λ = set(
            tsubtract ∘ Product((B_mva, n, T3), alfven_velocity_ani_ts);
            ylabel=ΔV_ylabel, labels=["Va*_L", "Va*_M", "Va*_N"]
        )
        push!(result, Λ, Va_Λ)
    end
    return result
end
```

@@

>>>>

Interactively exploring the quantities of solar wind. See [Data](000_data.qmd) for more details about different data products.

@@

```{julia}
using Dates
using Speasy
using DrWatson
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using Discontinuity
using SPEDAS
using DataFrames, DataFramesMeta
using GLMakie
using SPEDAS: set_meta, set_meta!
using PartialFunctions

import PSPAnisotropy.Labels as 𝑳
import PSPAnisotropy.YLabel as 𝒀

using Unitful
Unitful.preferunits(u"nA", u"km")
unitify(u) = (*)$u
unitify(u::String) = (*)$uparse(u)
```

## Event example (simple)

```{julia}
df, configs = workload()

psp_df = @rsubset(df, :id == "PSP"; view=true)
psp_config = configs[1]
wind_df = @rsubset(df, :id == "Wind"; view=true)
wind_config = configs[2]
thm_df = @rsubset(df, :id == "THEMIS"; view=true)
thm_config = configs[3]

thm_config["B"] = unitify("nT") ∘ thm_config["B"]
```

```{julia}
function plot_candidate(f, event, config, toffset=Second(0); kwargs...)
    tmin, tmax = event.t_us_ds
    t0 = tmin - toffset
    t1 = tmax + toffset

    B_product = config["B"]
    V_product = config["V"]

    B = B_product(t0, t1) |> set_meta!(; ylabel="B", labels=𝑳.B_RTN)
    V = V_product(t0, t1) |> set_meta!(; ylabel="V", labels=𝑳.V_RTN)
    n = config["n"](t0, t1)

    B_mva = mva(B, B) |> set_meta(; labels=𝑳.B_LMN)
    Va = Alfven_velocity_ts(B_mva, n) .|> u"km/s" |> set_meta(; labels=𝑳.Va_LMN, ylabel="V")
    V_mva = mva(V, B) |> set_meta(; labels=𝑳.V_LMN)
    J = hcat(current_density(B_mva, V_mva)...) |> set_meta(; labels=𝑳.J, ylabel="J")

    result = [tnorm_combine(B_mva), tsubtract(V_mva), n, tsubtract(Va), J]

    f = tplot(f, result; kwargs...)
    tlines!(f, [tmin, tmax])
    result
end

f = Figure(; size=(1200, 800))
res_psp = plot_candidate(f[1, 1], psp_df[18, :], psp_config, Second(10))
# res_thm = plot_candidate(f[1, 2], thm_df[10, :], thm_config)
# res_thm = plot_candidate(f[1, 2], thm_df[14, :], thm_config)
# res_thm = plot_candidate(f[1, 2], thm_df[17, :], thm_config)
# res_wind = plot_candidate(f[1, 3], wind_df[1, :], wind_config, Second(10))
f

@@

```{julia}
using Speasy
using SPEDAS: tnorm_combine, tplot, DataSet
using DimensionalData
using CairoMakie
# using GLMakie
using Statistics
using PartialFunctions
using Unitful
using Beforerr: _theme_legend, easy_save
using SPEDAS.SpaceDataModel: set, extend, @set
include("../src/psp_temp.jl")
include("../src/meta.jl")

import Discontinuity.AlgebraOfGraphics as AoG

update_theme!(;
    figure_padding=1,
    fonts=(; regular=AoG.firasans("Regular"), bold=AoG.firasans("Medium")),
    Legend=(; framevisible=false, padding=(0, 0, 0, -15)),
    Axis=(; xlabelfont=:bold, ylabelfont=:bold),
)

@@

```{julia}
psp_B = set(tnorm_combine ∘ PSP.B_1MIN; ylabel=𝒀.B, labels=𝑳.B_RTN)
psp_V = set(PSP.V; ylabel=𝒀.V, labels=𝑳.V_RTN)
psp_n = set(PSP.n_spi; ylabel=𝒀.n, yscale=identity)

tvars_psp = [psp_B, psp_n, psp_V]
tvars_psp_all = [
    psp_B,
    set(PSP.n; ylabel=𝒀.n),
    psp_V,
    set(PSP.T; ylabel=𝒀.T)
]
```

```{julia}
wind_B = set(tnorm_combine ∘ Wind.B_GSE_1MIN; ylabel=𝒀.B, labels=𝑳.B_XYZ)
wind_V = replace_outliers! ∘ set(Wind.V_GSE_K0; ylabel=𝒀.V, labels=𝑳.V_XYZ)

tvars_wind = [wind_B, set(Wind.n_p_K0; ylabel=𝒀.n), wind_V]
tvars_wind_all = [
    wind_B,
    set(Wind.n; ylabel=𝒀.n),
    wind_V,
    set(Wind.T; ylabel=𝒀.T)
]
```

@@

```{julia}
tvars_themis = [
    set(tnorm_combine ∘ THEMIS.B_GSE; ylabel=𝒀.B, labels=𝑳.B_XYZ),
    set(THEMIS.n; ylabel=𝒀.n),
    replace_outliers! ∘ set(THEMIS.V_GSE; ylabel=𝒀.V, labels=𝑳.V_XYZ),
    set(THEMIS.eTemp; ylabel=𝒀.T)
]
```

```{julia}
update_theme!(; Lines=(linewidth=1, alpha=0.9), figure_padding=(0, 0, 0, 15))

for enc in 7:9
    (tmin_psp, tmax_psp), (tmin_wind, tmax_wind) = get_timerange(enc)

@@

    legend = (; position=Top())
    labels = '1':'9'
    padding = (0, 30, 0, -30)

    f = Figure(; size=(800, 450))
    plots = tplot(f[1, 1], tvars_psp[1:3], tmin_psp, tmax_psp; legend)
    add_labels!(f[1, 1]; labels, open="(a.", padding)
    tplot(f[1, 2], tvars_wind[1:3], tmin_wind, tmax_wind; legend)
    add_labels!(f[1, 2]; labels, open="(b.", padding)
    # tplot(f[1, 3], _tmask! .∘ tvars_themis[1:3], tmin_wind, tmax_wind; legend)
    # add_labels!(f[1, 3]; labels, open="(c.")
    easy_save("overview-$enc")

    # f = Figure(; size=(1200, 600))
    # plots = tplot(f[1, 1], tvars_psp, tmin_psp, tmax_psp; legend)
    # add_labels!(f[1, 1]; labels, open="(a.")
    # tplot(f[1, 2], tvars_wind, tmin_wind, tmax_wind; legend)
    # add_labels!(f[1, 2]; labels, open="(b.")
    # tplot(f[1, 3], _tmask! .∘ tvars_themis, tmin_wind, tmax_wind; legend)
    # add_labels!(f[1, 3]; labels, open="(c.")
    # easy_save("overview-$enc-all")
end
```

@@

```{julia}
psp_B_mva = set(mva ∘ PSP.B; name="mva_B", labels=𝑳.B_LMN)
psp_V_mva = Product((PSP.V, PSP.B), mva_transform; ylabel=𝒀.V, labels=𝑳.V_LMN)
psp_eTemp_ani = Product(read_electron_temperature(), tview; labels=["Electron (parallel)", "Electron (perpendicular)"])
psp_pTemp_ani = Product(read_proton_temperature(), tview; labels=["Proton (parallel)", "Proton (perpendicular)"])
psp_Temp_ani = [psp_eTemp_ani, psp_pTemp_ani]
psp_mva_all = [psp_B_mva, psp_V_mva, psp_Temp_ani]

thm_B_mva = set(mva ∘ THEMIS.B_FGL_GSE; name="mva_B", labels=𝑳.B_LMN)
thm_V_mva = Product((THEMIS.V_GSE, THEMIS.B_FGL_GSE), mva_transform; ylabel=𝒀.V, labels=𝑳.V_LMN)
thm_mva_all = [thm_B_mva, thm_V_mva, THEMIS.T_ani]
```

@@

## Thickness, current density comparison and Alfvénicity

```{julia}
using AlgebraOfGraphics
using Discontinuity: compute_params!, filter_low_mva_quality, compute_Alfvenicity_params!
using Beforerr
using Beforerr: add_labels!
using CairoMakie
using StatsBase
CairoMakie.activate!()

set_aog_theme!()
formats = (:pdf,)

foreach([:log10, :log2, :log]) do f
    @eval import Base: $f
    @eval $f(x::Unitful.Quantity) = $f(ustrip(x))
end

df, configs = workload()
𝐧 = :mva
𝒎 = var_mapping(; 𝐧)

compute_params!(df)

# AoG fix
using LaTeXStrings
import AlgebraOfGraphics: append_unit_string
append_unit_string(s::LaTeXStrings.LaTeXString, u::String) = s

subset_𝐧(df, 𝐧) = 𝐧 == :mva ? filter_low_mva_quality(df) : df

function alpha_layer(df, alpha)
    psp_df = @rsubset(df, :id == "PSP"; view=true)
    thm_df = @rsubset(df, :id == "THEMIS"; view=true)
    wind_df = @rsubset(df, :id == "Wind"; view=true)
    (data(thm_df) + data(wind_df)) * visual(; alpha) + data(psp_df)
end
```

```{julia}
fname = "properties_xy_comp-enc"

for enc in 7:9, 𝐧 in [:mva, :cross]
    view = true
    𝒎 = var_mapping(; 𝐧)

    tdf = subset_𝐧(dropmissing(@subset(df, :enc .== enc; view)), 𝐧)
    compute_Alfvenicity_params!(tdf)
    plt = alpha_layer(tdf, 0.618) * visual(Scatter; markersize=8, legend=(; alpha=1)) * mapping(color=:id, marker=:id)

    fig = Figure(size=(800, 300))
    axis = (; yscale=log10, xscale=log10)
    v_axis = (; yscale=log10, xscale=log10, limits=((2, 400), (2, 400)))

    ga = fig[1, 1]
    gb = fig[1, 2]
    gc = fig[1, 3]

    spec1 = plt * mapping(𝒎.l, 𝒎.j)
    spec2 = plt * mapping(𝒎.l_norm, 𝒎.j_norm)
    spec3 = plt * mapping(𝒎.v_Alfven, 𝒎.v_ion)

    grid1 = draw!(ga, spec1; axis)
    draw!(gb, spec2; axis)
    draw!(gc, spec3; axis=v_axis)

    slopes = (1.0, 0.7, 0.4, 0.1)
    foreach(slopes) do s
        lines!(1 .. 1000, (*)$s, linestyle=:dash, label="$s")
    end
    axislegend("slope", position=:lt)

    pretty_legend!(fig, grid1)
    add_labels!([ga, gb, gc])
    easy_save("$fname-$enc-$𝐧"; formats, force=true)
    display(fig)
end
```

![](../figures/properties_xy_comp-enc-7-mva.pdf)
![](../figures/properties_xy_comp-enc-7-cross.pdf)
![](../figures/properties_xy_comp-enc-8-mva.pdf)
![](../figures/properties_xy_comp-enc-8-cross.pdf)
![](../figures/properties_xy_comp-enc-9-mva.pdf)
![](../figures/properties_xy_comp-enc-9-cross.pdf)

```{julia}
fname = "properties_hist-enc"

for enc in 7:9, 𝐧 in [:mva, :cross]
    view = true
    𝒎 = var_mapping(; 𝐧)
    datalimits_f = x -> quantile(x, [0.02, 0.98])
    # drop missing and NaN
    tdf = @chain df begin
        @subset(:enc .== enc, .!isnan.(:L_n_cross_norm); view)
        dropmissing
        subset_𝐧(𝐧)
        compute_Alfvenicity_params!
        @subset(.!isnan.(:v_l_ratio))
    end

    spec = data(tdf) *
           mapping([𝒎.l_log, 𝒎.j_log, 𝒎.l_norm_log, 𝒎.j_norm_log, 𝒎.v_l_ratio, 𝒎.bn], layout=AoG.dims(1) => renamer(ones(String, 6))) *
           mapping(color=:id) *
           AoG.density(datalimits=datalimits_f)

    draw(spec, scales(Layout=(; palette=wrapped(cols=2))); facet=(; linkyaxes=:none))
    easy_save("$fname-$enc-$𝐧"; formats, force=true)
end
```

```{julia}
for enc in 7:9
    tdf = filter_low_mva_quality(dropmissing(@subset(df, :enc .== enc; view=true), view=true))
    data_layer = data(tdf)
    legend = (; markersize=10, alpha=1)
    plt = data_layer * visual(Scatter; markersize=4, alpha=0.618, legend) * mapping(color=:id, marker=:id)

    fig = Figure(size=(800, 400))
    axis = (; yscale=identity, xscale=identity)

    ga = fig[1, 1]
    gb = fig[1, 2]

    spec1 = plt * mapping(𝒎.bn, 𝒎.ω)
    spec2 = plt * mapping(𝒎.bn, 𝒎.ω_in)

    grid1 = draw!(ga, spec1; axis)
    draw!(gb, spec2; axis)

    pretty_legend!(fig, grid1)
    add_labels!([ga, gb])
    display(fig)
end
```

### 2025-06-12 notes.qmd
update

>>>>

# Research notes

## In-situ synergistic observations (Alignment)

**Similar** type of solar wind or **Same** solar solar wind

How to define alignment?

- Trajectory & Orbits (theory): ballistic approximation of Parker spiral

- Plasma properties (observation):

  - Invariants: plasma composition, ionization states, mass flux $n u r^2$

  - Correlation: magnetic field / plasma velocity

- Simulation: Trace plasma parcels evolution

## Anisotropy

### Estimated Anisotropy

@shenComparingPlasmaAnisotropy2024

The equation for $Δv_A$ includes a factor depending on the anisotropy of the plasma.

Pressure anisotropy $Λ = \mu (P_∥ - P_⊥) / B^2$.

![](figures/enc7/anisotropy_theory.png)

### Comparison between Estimated and Observed Anisotropy

![](figures/enc7/anisotropy_comparison_2d.png)
